Trending assessment for nowcasts, measurements, and forecasts

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

\todo{Signal/Prediction vereinheitlichen}
\subsection{Basics of trending and four-quadrant plots}\label{subsec:trending-basics}

@@

All measures can also be evaluated as a rolling estimate to detect changes in performance over time.
A rolling estimate is a sequence of estimates, each computed on a length-$w$-window of the data.
For the trending ratio, a rolling estimate with a backward-looking window is given by
\begin{equation*}
    \acc_t (\diffx, \diffy) \coloneqq \frac{\sum_{t^* = t-w + 1}^{t} \ind{\diffxt[t^*] \diffyt[t^*] > 0}}{w} \quad t = w-1, \dots, T.\label{eq:acc_rolling}
\end{equation*}
Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
Thus, the yearly course of the prediction's trending ratio can be detected.
The trending ability of the signal has a strong yearly, sinus-shaped seasonality with higher trending ability peak after a quarter of a year and a low point after three quarters.
This seasonal behavior cannot be observed in the two four-quadrant plots as the high number of points overlays it or in the trending ratio.

@@

\todo{Übergang zu Bootstrapping}
Confidence intervals can account for the estimation uncertainty of the measures above.
Bootstrap confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
New samples are drawn with replacement from the dataset in bootstrapping and are not based on parametric assumptions as for classical confidence intervals.
The confidence interval is then computed based on the bootstrap samples.
We examine three methods for bootstrapping here: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
In the \textit{percentile} approach, the confidence interval for level $\alpha$ is built directly from the bootstrap estimators' empirical distribution.
The \textit{basic} approach computes the confidence interval based on the non-bootstrapped estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but needs smaller datasets for reasonable confidence intervals.
The basic approach balances those two aims.
We compare the approaches in a small synthetic data study concerning their small-dataset behavior and computation time in Appendix~\ref{sec:appendix}.
For small datasets, the \ac{bca} holds the confidence level also for small data sets with a negligibly higher computation time than the other approaches.
Thus, we use the \ac{bca} method for the confidence intervals in the applications in Section~\ref{sec:application}.

>>>>

\subsection{Basics of trending and four-quadrant plots}\label{subsec:trending-basics}

@@

All measures can also be evaluated as a rolling estimate to detect changes in performance over time.
A rolling estimate is a sequence of estimates, each computed on a length-$w$-window of the data.
For the trending ratio, a rolling estimate with a backward-looking window is given by
\begin{equation*}
    \acc_t (\diffx, \diffy) \coloneqq \frac{\sum_{t^* = t-w + 1}^{t} \ind{\diffxt[t^*] \diffyt[t^*] > 0}}{w} \quad t = w-1, \dots, T.\label{eq:acc_rolling}
\end{equation*}
Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
Thus, the yearly course of the prediction's trending ratio can be detected.
The trending ability of the signal has a strong yearly, sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
This seasonal behavior cannot be observed in the two four-quadrant plots as the high number of points overlays it or in the trending ratio.

@@

Confidence intervals can account for the estimation uncertainty of the measures above.
Bootstrap confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
New samples are drawn with replacement from the dataset in bootstrapping and are not based on parametric assumptions as for classical confidence intervals.
The confidence interval is then computed based on the bootstrap samples.
We examine three methods for bootstrapping here: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
In the \textit{percentile} approach, the confidence interval for level $\alpha$ is built directly from the bootstrap estimators' empirical distribution.
The \textit{basic} approach computes the confidence interval based on the non-bootstrapped estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but needs smaller datasets for reasonable confidence intervals.
The basic approach balances those two aims.
We compare the approaches in a small synthetic data study concerning their small-dataset behavior and computation time in Appendix~\ref{sec:appendix}.
For small datasets, the \ac{bca} holds the confidence level also for small data sets with a negligibly higher computation time than the other approaches.
Thus, we use the \ac{bca} method for the confidence intervals in the applications in Section~\ref{sec:application}.
